
\begin{thebibliography}{12}

\bibitem{clrs}
Charles E. Leiserson, Thomas H. Cormen, Clifford Stein, Ronald Rivest.
\textit{Introduction to Algorithms}, MIT Press, 1990

\bibitem{neel}
David L. Neel, Nancy Ann Neudauer. \textit{Matroids You Have Known}\\
\texttt{http://www.maa.org/sites/default/files/pdf/shortcourse/2011/matroidsknown.pdf}

\bibitem{geoff}
Geoff Whittle. \textit{Rota's Conjecture}\\
\texttt{http://www.asiapacific-mathnews.com/04/0401/0011\_0012.pdf}

\bibitem{matroidtheory}
James Oxley. \textit{Matroid Theory}, Oxford University Press, 2011

\bibitem{oxley}
James Oxley. \textit{WHAT IS A MATROID?}\\
\texttt{https://www.math.lsu.edu/\~{}oxley/survey4.pdf}

\bibitem{jayant}
Jayant Apte. \textit{Solving Rota's Conjecture*}\\ 
\texttt{http://www.ece.drexel.edu/walsh/Jayant\_RotaConjecture1.pdf}

\bibitem{gf4}
Jim Geelen. \textit{The Excluded Minors for GF(4)-Representable Matroids}\\
\texttt{https://www.math.uwaterloo.ca/\~{}jfgeelen/Publications/gf4.pdf}

\bibitem{solving}
Jim Geelen, Bert Gerards, Geoff Whittle. \textit{Solving Rota's Conjecture}\\
\texttt{http://www.ams.org/notices/201407/rnoti-p736.pdf}

\bibitem{lec9}
Michel X. Goemans, \textit{Lecture 9 from Advanced Combinatorial Optimization}\\
\texttt{http://math.mit.edu/\~{}goemans/18438F09/lec9.pdf}

\bibitem{hlineny}
Petr Hlin\v{e}n\'{y}. \textit{On Matroid Representability and Minor Problems}\\
\texttt{http://www.fi.muni.cz/\~{}hlineny/papers/mrepres-mfcs.pdf}

%\bibitem{rudi}
%"Advanced Graph Theory, lecture 1" Rudi Pendavingh

\end{thebibliography}
