\section{Introduction to Matroid}

A matroid is a mathematical structure that abstracts the concept of linear independence.
A linearly independent set of vectors has a lot of good properties.
Interestingly, a lot of sets of other mathematical objects often share those properties.
For example, any subset of a linearly independent set of vectors is always linearly independent.
Consider a subset of edges of a graph that does not contain any cycle.
Then any subset of it does not contain any cycle.
The matroid theory is an attempt to mathematically formalize those properties.

\begin{defn}
A matroid $M = (E, \mathcal{I})$ is a pair such that $E$ is a finite set of elements, and $\mathcal{I}$ is a family of subsets of $E$ with the following properties:
\begin{itemize}
\item $\emptyset \in \mathcal{I}$
\item For any $A\in \mathcal{I}$, any subset of $A$ is in $\mathcal{I}$.
\item For any $A, B \in \mathcal{I}$ such that $\lvert A \rvert < \lvert B \rvert$, there always exists $x \in B - A$ such that $A \cup \{ x \} \in \mathcal{I}$.
\end{itemize}
\end{defn}
$E$ is called the ground set, $\mathcal{I}$ is called a family of independent sets. A subset of $E$ is called independent if and only if it is in $\mathcal{I}$.
In this paper, we will put our focus on a matroid with a finite ground set. 


There are some basic matroids that are important in the following discussion.
We will start by introducing a column matroid. 
A column matroid is constructed from a matrix over a field $\mathbb{F}$.

\begin{defn}
Let a matrix $A$ with $m$ rows over $\mathbb{F}$ be given.
A column matroid $M$ of $A$ is a matroid with a ground set $\{ 1, 2, \cdots m \}$.
A subset of $E$ is independent in $M$ if and only if the set of the corresponding column vectors is linearly independent.
\end{defn}

\begin{thm}
A column matroid is indeed a matroid.
\end{thm}

\begin{proof}
First, an empty set of vectors is linearly independent by definition.
Suppose there exists a linearly independent set of vectors that has a linearly dependent subset.
Let $A = \{a_1, \cdots, a_k\} \subseteq B = \{b_1, \cdots, b_n\}$ be such sets.
Since $A$ is linearly dependent, there exist constants $c_1, \cdots, c_k$ such that $c_1 a_1 + \cdots + c_k a_k = 0$ and not all $c_i$'s are $0$.
However, that implies that we can find constants $d_1, \cdots, d_n$ such that $d_1 b_1 + \cdots + d_k b_k = 0$ and not all $d_i$'s are $0$.
This is a contradiction since $B$ is supposed to be independent.
Therefore, such a subset must not exist, and thus all subsets of linearly independent sets are linearly independent.
Now we want to prove the third property.
Let $A = \{a_1, \cdots, a_k\}, B = \{b_1, \cdots, b_n\}$ be linearly independent sets, and assume that $k < n$.
Suppose for each $i = 1, \cdots, n$, $b_i \in span\{ a_1, \cdots, a_k \}$.
It means that the span of $B$ is a subspace of the span of $A$, which has a smaller dimension than $n$.
That is a contradiction.
Therefore, there exists $i$ such that $b_i \notin span \{ a_1, \cdots, a_k \}$.
For such $i$, $\{ a_1, \cdots, a_k, b_i \}$ should be linearly independent.
Since a column matroid satisfies the three properties, it is indeed a matroid.
\end{proof}


\begin{defn}
A uniform matroid $U_{r, k}$ is a matroid such that $E = \{ 1, \cdots, k \}$ and $\mathcal{I} = \{ X \mid  X \subseteq E, \lvert X \rvert \leq r \}$.
\end{defn}

It is easy to see that a uniform matroid is indeed a matroid.

% If I need more pages, just introduce a graph matroid

Here are some important results that will show up later in this paper.

\begin{thm}
All maximal independent sets have the same size.
\end{thm}

\begin{proof}
Let $X, Y$ be maximal independent sets of some matroid.
Suppose $\lvert X \rvert \neq \lvert Y \rvert$.
Without loss of generality, $\lvert X \rvert < \lvert Y \rvert$.
By the third property of a matroid, there exists $e \in Y - X$ such that $X \cup \{ e \}$ is independent.
It is a contradiction since $X$ is a maximal independent set.
Therefore $\lvert X \rvert = \lvert Y \rvert$.
\end{proof}

This is indeed true in linear algebra.
Given a matrix, any maximal independent subset of column vectors always has the same size. In linear algebra, this number is often referred to as the \textit{dimension} of a vector space or the \textit{rank} of a matrix.
In the matroid theory, we use the term \textit{rank} as well.


\begin{defn}
The rank of a matroid is the size of a maximal independent set.
\end{defn}

The rank of a column matroid is equal to the rank of the matrix since the subset of a ground set is independent if and only if the subset of column vectors is linearly independent.

\begin{defn}
Let $M = (E, \mathcal{I})$ be given.
$e \in E$ is called a loop if $\{ e \} \notin \mathcal{I}$.
\end{defn}

We will conclude this chapter by introducing the notion of isomorphic matroids.
\begin{defn}
Let $M_1 = (E_1, \mathcal{I}_1), M_2 = (E_2, \mathcal{I}_2)$ be given.
$M_1, M_2$ are isomorphic to each other if there exists a bijective mapping $\phi: E_1 \rightarrow E_2$ such that
$\forall X \subseteq E_1, X \in \mathcal{I}_1 \iff \{ \phi(e) : e \in X \} \in \mathcal{I}_2$.
\end{defn}

Two isomorphic matroids have the same structure.
